---
title: "Linking Tariffline HS with FCL"
author: "Alex Matrunich"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Linking Tariffline HS with FCL}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r set_options, echo = F, warning = F}
suppressPackageStartupMessages(library(dplyr))

knitr::opts_chunk$set(list(echo = FALSE,
                           eval = TRUE,
                           cache = TRUE,
                           warning = FALSE,
                           message = FALSE,
                           fig.height = 10,
                           fig.width = 12))

source(file.path(Sys.getenv("HOME"), ".pwd.R"))

source(file.path(Sys.getenv("HOME"), 
                 "r_adhoc", 
                 "trade_prevalid_testing", 
                 "setupconnection.R"))

trade_src <- src_postgres("sws_data", "localhost", 5432, "trade", .pwd, 
                          options = "-c search_path=ess")

agri_db <- tbl(trade_src, sql("
select * from ess.agri
"))
```

1. Convert country codes.
2. Align HS-codes as numbers (max length among Tariffline, fromcode and tocode).

# Country codes

We use countrycode package.

```{r countrycode}
agri_db %>% 
  filter(year == "2011") 
```

All reporters' codes in Tariffline are presented in ComTrade country code scheme:

```{r}
ctm49 <- getComtradeM49()$code
unm49 <- getOfficialM49()$code

tlreps <- agri_db %>% 
  select(reporter) %>% 
  distinct() %>% 
  collect()

length(tlreps$reporter[!is.element(tlreps$reporter, ctm49)])
```

It is more reliable to make conversions through the names of countries rather than digital codes. We can not use FAOSTAT package, as there is no codes from ComTrade.

```{r}
tlreps <- tlreps %>% 
  left_join(getComtradeM49(),
            c("reporter" = "code"))
```

We got country names. Now we try to match them to FAO area codes using reqular expressions from countrycode package.

```{r, warning = T}
tlreps <- tlreps %>% 
  mutate(faoarea = countrycode::countrycode(name, "country.name", "fao", 
                                            warn = T))

tlreps %>% 
  filter(is.na(faoarea))

```

The problem with Palestine, for example, that countrycode() does not ignore case during reqular expressions matching.

```{r}
tlreps$faoarea[tlreps$name == "State of Palestine"] <- 299
tlreps$faoarea[tlreps$name == "New Caledonia"] <- 153 # No code for FAO in countrycode
tlreps$faoarea[tlreps$name == "Greenland"] <- 85 # No code for FAO in countrycode
tlreps$faoarea[tlreps$name == "Mayotte"] <- 270 # No code for FAO in countrycode
tlreps$faoarea[tlreps$name == "Bermuda"] <- 17 # No code for FAO in countrycode
tlreps$faoarea[tlreps$name == "French Polynesia"] <- 70 # No code for FAO in countrycode
tlreps$faoarea[tlreps$name == "Turks and Caicos Isds"] <- 224 # No code for FAO in countrycode
tlreps$faoarea[tlreps$name == "Aruba"] <- 22 # No code for FAO in countrycode
tlreps$faoarea[tlreps$name == "Montserrat"] <- 142 # No code for FAO in countrycode
tlreps$faoarea[tlreps$name == "China, Hong Kong SAR"] <- 96 # No code for FAO in countrycode
tlreps$faoarea[tlreps$name == "China, Macao SAR"] <- 128 # No code for FAO in countrycode
tlreps$faoarea[tlreps$name == "China"] <- 41 # In countrycode it is linked to China including Hong Kong and Macao

  
```

## Partners


```{r}
tlprts <- agri_db %>% 
  select(partner) %>% 
  distinct() %>% 
  collect()

tarifflineData <- agri_db %>% 
  filter(year == "2011",
         flow == "1") %>% 
  group_by(reporter, partner, hs2) %>% 
  summarise(value = sum(value)) %>% 
  collect() %>% 
  mutate(value = round(value, 0)) %>% 
  ungroup()

# comtradeData <- getCommodityFullData(c("01", "02"), "5621", "2011") %>% 
comtradeDataRaw <- getCommodityFullData(getAgriHSCodes(), c("5621"), "2011")  

comtradeData <- comtradeDataRaw %>% 
  select(reporter = starts_with("reporting"),
         ct_partner  = starts_with("partner"),
         hs      = ends_with("HS"),
         ct_value    = Value) %>% 
  mutate(hs2 = stringr::str_extract(hs, "^.{2}")) %>% 
  group_by(reporter, ct_partner, hs2) %>% 
  summarize(ct_value = round(sum(ct_value, na.rm = T), 0)) %>% 
  ungroup()
  

joinedByValue <- tarifflineData %>% 
  left_join(comtradeData,
            by = c("reporter", "hs2", "value" = "ct_value"))


joinedByValue %>% 
  filter(!is.na(ct_partner)) %>% 
  select(reporter, partner, ct_partner) %>% 
  group_by(reporter, partner) %>% 
  mutate(n = n(),
         matches = unique(partner) %in% ct_partner,
         matchesminus = unique(partner) %in% as.character(as.integer(ct_partner) - 1)) %>% 
  ungroup() %>% 
  filter(!matches,
         !matchesminus,
         partner != "840") %>% 
  arrange(desc(n)) %>% 
  as.data.frame()
  

joinedByValue %>% 
  filter(is.na(ct_partner))

partlyFinished <- joinedByValue %>% 
  select(reporter, partner, ct_partner) %>% 
  group_by(reporter, partner) %>% 
  mutate(variants = length(unique(ct_partner)),
         matches = unique(partner) %in% ct_partner,
         matchesminus1 = unique(partner) %in% as.character(as.integer(ct_partner) - 1),
         matchesminus2 = unique(partner) %in% as.character(as.integer(ct_partner) - 2)) %>% 
  ungroup() %>% 
  mutate(correct_var = ifelse(matches,
                              partner,
                              ifelse(matchesminus1, 
                                     as.character(as.integer(ct_partner) - 1),
                                     ifelse(matchesminus2, 
                                            as.character(as.integer(ct_partner) - 2),
                                            NA))))


finishedCodes <- partlyFinished %>% 
  select(reporter, partner, correct_var) %>% 
  filter(!is.na(correct_var)) %>% 
  distinct() 
  
x <- joinedByValue %>% 
  select(reporter, partner, ct_partner) %>% 
  group_by(reporter, partner, ct_partner) %>% 
  mutate(ranking = n()) %>% 
  group_by(reporter, partner) %>% 
  filter(ranking == max(ranking)) %>% 
  distinct() %>% 
  filter(ct_partner != partner &
           partner != as.character(as.integer(ct_partner) - 1) &
           partner != as.character(as.integer(ct_partner) - 2)) %>% 
  as.data.frame() %>% 
    arrange(partner) 
# So there are 78 cases when reporter-specific partner's code doesn't match Comtrade's M49

# Codes that are absent in Comtrade list 
unique(x$ct_partner[!is.element(x$ct_partner, getComtradeM49()$code)])

# Codes that are absent in SWS list 
unique(x$ct_partner[!is.element(x$ct_partner, getAllReportersRaw()$code)])
```

```{r}
partlyFinished %>% 
  group_by(reporter, partner, ct_partner, correct_var) %>% 
  summarize(ranking = n()) %>% 
  group_by(reporter, partner) %>% 
  mutate(varavail = any(!is.na(correct_var)) | any(!is.na(ct_partner))) %>% 
  filter(!varavail) %>% 
  filter(!is.element(partner, getAllCountryCodes())) %>% 
  select(partner) %>% 
  

```

666 cases, where partner is not available


# By HS6

```{r}

tarifflineData <- agri_db %>% 
  filter(year == "2011",
         flow == "1") %>% 
  group_by(reporter, partner, hs6) %>% 
  summarise(value = sum(value)) %>% 
  collect() %>% 
  mutate(value = round(value, 0)) %>% 
  ungroup()

comtradeData <- comtradeDataRaw %>% 
  select(reporter = starts_with("reporting"),
         ct_partner  = starts_with("partner"),
         hs6      = ends_with("HS"),
         ct_value    = Value) 
  

joinedByValue <- tarifflineData %>% 
  left_join(comtradeData,
            by = c("reporter", "hs6", "value" = "ct_value"))

partlyFinished <- joinedByValue %>% 
  select(reporter, partner, ct_partner) %>% 
  group_by(reporter, partner) %>% 
  mutate(matches = unique(partner) %in% ct_partner,
         matchesminus1 = unique(partner) %in% as.character(as.integer(ct_partner) - 1),
         matchesminus2 = unique(partner) %in% as.character(as.integer(ct_partner) - 2)) %>% 
  ungroup() %>% 
  mutate(correct_var = ifelse(matches,
                              partner,
                              ifelse(matchesminus1, 
                                     as.character(as.integer(ct_partner) - 1),
                                     ifelse(matchesminus2, 
                                            as.character(as.integer(ct_partner) - 2),
                                            NA))))

partlyFinished %>% 
  group_by(reporter, partner, ct_partner, correct_var) %>% 
  summarize(ranking = n()) %>% 
  group_by(reporter, partner) %>% 
  mutate(varavail = any(!is.na(correct_var)) | any(!is.na(ct_partner))) %>% 
  ungroup() %>% 
  filter(!varavail) %>% 
  select(partner) %>% 
  distinct() %>% 
  # filter(!is.element(partner, getAllCountryCodes()))
  filter(!is.element(partner, getOfficialM49()$code))

```

### Partner reference list

```{r}
unsdlist <- XLConnect::readWorksheetFromFile(system.file("extdata",
                                                         "Partner reference list.xlsx",
                                                         package = "tradeproc"),
                                             sheet = "countryList")

notInList <- tlprts$partner[!is.element(tlprts$partner, unsdlist$crCode)]

agri_db %>% 
  filter(year == "2011",
         partner %in% notInList)
```

